---
title: "Case Study 2"
author: "Jeff Johnson"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=7, fig.width=12, fig.align="center")
```

## A Case Study on Attrition and other measures

Greetings and please find in this document the means to predict monthly income and attrition with some additional analysis as well.

```{r}


#load libraries
library(readr)
library(ggplot2)
library(plotly)
library(GGally)
library(class)
library(caret)
library(dplyr)
library(e1071)

#Load data
BusinessData <- read.csv(file.choose())

#Modify some columns for usage in models
BusinessData$Attrition = as.factor(BusinessData$Attrition)
BusinessData$MonthlyIncome = as.numeric(BusinessData$MonthlyIncome)
BusinessData$MonthlyRate = as.numeric(BusinessData$MonthlyRate)
BusinessData$TotalWorkingYears = as.numeric(BusinessData$TotalWorkingYears)
BusinessData$PercentSalaryHike = as.numeric(BusinessData$PercentSalaryHike)
BusinessData$OverTime = as.factor(BusinessData$OverTime)
```

## Some Useful Insights

Here I have included some insights from the data.

```{r}
# show count of those that have left the company by job role
BusinessData %>% filter(Attrition == "Yes") %>% group_by(JobRole) %>% summarise(count = n())

#plot attrition by job role with those already gone
p1 <- BusinessData %>% filter(Attrition == "Yes") %>% ggplot(aes(x = JobRole, fill = JobRole)) + geom_bar() + ggtitle("Attrition by Job Role")
ggplotly(p1)

```
The Job Roles with the highest attrition are Sales Executive, Research Scientist and Laboratory Technician. The lowest are Research Director, Manufacturing Director and Manager.

```{r}
# plot attrition by job role with those both here and gone
p2 <- BusinessData %>% ggplot(aes(x = JobRole, fill = Attrition)) + geom_bar() + ggtitle("Attrition by Job Role")
ggplotly(p2)

```
We can see in the above plot the percent of those that have left and chosen to stay.


```{r}
# plot Attrition by Gender with those that have already left
p3 <- BusinessData %>% filter(Attrition == "Yes") %>% ggplot(aes(x = Gender, fill = Gender)) + geom_bar() + ggtitle("Attrition by Gender")
ggplotly(p3)

```
Attrition is highest in males, with 62% of those in the dataset leaving the company being Male.

```{r}
# plot for Monthly Income by Job Role featuring Attrition
BusinessData %>% ggplot(aes(x = MonthlyIncome, y = JobRole, color = Attrition)) + geom_boxplot() + ggtitle("Monthly Income by Job Role and Attrition")

```
We can see here that those in the lower range of monthly income might be leaving due to a discrepancy in pay.


```{r}
# plot for overtime employees to show high percentage of overtime employees having left
p4 <- BusinessData %>% ggplot(aes(x = OverTime, fill = Attrition)) + geom_bar() + ggtitle("Attrition by Exempt and Non-Exempt Employees")
ggplotly(p4)


```
One of the 3 contributing factors to attrition I've found is those that are non-exempt are leaving at a higher rate.


```{r}
# plot monthly income vs total working years featuring attrition
BusinessData %>% ggplot(aes(x = MonthlyIncome, y = TotalWorkingYears, color = Attrition)) + geom_point() + geom_smooth(method = 'lm', linetype = "dashed", color = "darkred", fill="blue") + ggtitle("Monthly Income vs Total Working Years with Attrition")

```
A positive correlation can be seen here between Monthly Income and Total Working Years, with a linear model thrown in.

## Training and Test Set

Here is the code for specifying the training and test sets to be used in the models (a 70/30 split).
```{r}

set.seed(6)
sample_size=floor(0.7 * nrow(BusinessData)) #set sample size for training indices
trainIndices=sample(seq_len(nrow(BusinessData)), size = sample_size) #set training indices for training set and inversely test
attrition_train = BusinessData[trainIndices,] #create training set
attrition_test = BusinessData[-trainIndices,] #create test set

```

## k-NN model

Here is the k-NN model
```{r}

#k-NN model with k as square root of number of observations in training set
#attrition_knn <- knn(attrition_train[,c(18, 20, 24)],attrition_test[,c(18, 20, 24)],attrition_train$Attrition,k=sqrt(nrow(attrition_train)))

#confusionMatrix(attrition_knn,as.factor(attrition_test$Attrition)) #confusion matrix to show statistics on model

```

## Naive Bayes model

Here is the Naive Bayes model
```{r}
# set naive bayes model
attrition_nb <- naiveBayes(attrition_train[, c(18, 20, 24)], attrition_train$Attrition) #set Naive Bayes model with training set

confusionMatrix(table(predict(attrition_nb, attrition_test[, c(18, 20, 24)]), attrition_test$Attrition)) #confusion matrix to show statistics on model

```

## Linear model

Here is the Linear model, based on Monthly Income, Total Working Years and Percent Salary Hike (Multiple Linear Regression)
```{r}

salary_lm_model = lm(MonthlyIncome ~ TotalWorkingYears + PercentSalaryHike ,data=BusinessData) #set linear model
lm_predict <- predict(salary_lm_model, BusinessData)

difference <- lm_predict - BusinessData$MonthlyIncome

#compute and show RMSE
rmse <- 12*sqrt(mean(difference^2))
rmse
summary(salary_lm_model)

confint(salary_lm_model, conf.level=0.95)

#plot the model
plot(salary_lm_model)


```